---
title: "Interpret parameters of interactive models in REFS"
output: html_document
---

In Roche projects, we fitted interactive effects between treatment arms and other variables. However, interpretations of the model parameters are not straight forward. This document aims to answer this question. 

Check model parameters when fitting interactive models in R, before inspecting REFS otuputs.

```{r}

library(dplyr)
library(contrast)

# interaction between two binary variables
x1 = sample(c(0, 1), 100, replace = T)
x2 = sample(c(0, 1), 100, replace = T)
# response as continuous
y1 = rnorm(100, 1, 1)

# summarize the response (y) by predictor conditions
(y1.avg = aggregate(y1 ~ x1 + x2, FUN = "mean"))

# fit interactive model
summary(glm(y1 ~ x1 * x2))

# parameter interpretation
# intercept: expected y when both x1 and x2 equal 0 (basal level)
# x1: expected y when x1 equals 1 and x2 equals 0, conditional on the intercept
y1.avg[y1.avg$x1 == 1 & y1.avg$x2 == 0, "y1"] - y1.avg[y1.avg$x1 == 0 & y1.avg$x2 == 0, "y1"]
# x2: expected y when x1 equals 0 and x2 equals 1, conditional on the intercept
y1.avg[y1.avg$x1 == 0 & y1.avg$x2 == 1, "y1"] - y1.avg[y1.avg$x1 == 0 & y1.avg$x2 == 0, "y1"]
# x1:x2: expected y when x1 equals 1 and x2 equals 1, conditional on intercept, x1, and x2
y1.avg[y1.avg$x1 == 1 & y1.avg$x2 == 1, "y1"] - y1.avg[y1.avg$x1 == 0 & y1.avg$x2 == 0, "y1"] -
  (y1.avg[y1.avg$x1 == 0 & y1.avg$x2 == 1, "y1"] - y1.avg[y1.avg$x1 == 0 & y1.avg$x2 == 0, "y1"]) -
  (y1.avg[y1.avg$x1 == 1 & y1.avg$x2 == 0, "y1"] - y1.avg[y1.avg$x1 == 0 & y1.avg$x2 == 0, "y1"])

```

Note: 
1. R output for interactive models are consistent, regardless of the data type of response and predictor variables
2. we use two binary predictor and continuous response because it is the easist to interpret
3. but the interactive parameter is still confusing, especially for the interactive term
4. and people usually construct more meaningful constrast taking R::lm outputs 

REFS output

```{r}
library(REFSfs)
library(survival)

# this is starmaster
ens = fsReadModel("/users/xwang/roche34/otf2/m1.new/NetFragFile.txt.protobuf")

# extract single network
net = fsSubsetEnsemble(ens, 1)

# netfrag parameters of the network
otf = fsGetFrags(net, "pfs_AVAL")
otf$formula

# compare with R output to understand meanings of OTF parameters 

load("~/roche34/otf2/alldata.rdt")
alldata = refsdf.c1
vars = fsCausalVars(net, "pfs_AVAL", maxpath = 1)

load("~/roche34/otf2/alldata.rdt")
mydf = refsdf.c1[c(vars, "pfs_AVAL", "pfs_CNSR")]
mydf$pfs = Surv(mydf$pfs_AVAL, mydf$pfs_CNSR)
mydf$asl.treat.char_ARMCD

fit = survreg(pfs ~ asl.treat.char_ARMCD * aig_IGM + 
                    asl.treat.char_ARMCD * alb_Albumin +
                    asl.treat.char_ARMCD * alb_Platelet +
                    asl.treat.char_ARMCD * asl.clin.num_EXINV2 +
                    asl.treat.char_ARMCD * asl.lab.num_LDH1,
                    data = mydf)

coef(fit)

length(coef(fit))

# interactive models can be described by making dummy variables

mydf$d0 = as.numeric(mydf$asl.treat.char_ARMCD == 0) 
mydf$d1 = as.numeric(mydf$asl.treat.char_ARMCD == 1) 
mydf$d0_aig_IGM = mydf$d0 * mydf$aig_IGM
mydf$d1_aig_IGM = mydf$d1 * mydf$aig_IGM
mydf$d0_alb_Albumin = mydf$d0 * mydf$alb_Albumin
mydf$d1_alb_Albumin = mydf$d1 * mydf$alb_Albumin
mydf$d0_alb_Platelet = mydf$d0 * mydf$alb_Platelet
mydf$d1_alb_Platelet = mydf$d1 * mydf$alb_Platelet
mydf$d0_asl.clin.num_EXINV2 = mydf$d0 * mydf$asl.clin.num_EXINV2
mydf$d1_asl.clin.num_EXINV2 = mydf$d1 * mydf$asl.clin.num_EXINV2
mydf$d0_asl.lab.num_LDH1 = mydf$d0 * mydf$asl.lab.num_LDH1
mydf$d1_asl.lab.num_LDH1 = mydf$d1 * mydf$asl.lab.num_LDH1

fit = survreg(pfs ~ d0 + d1 + 
                    d0_aig_IGM + d1_aig_IGM +  
                    d0_alb_Albumin + d1_alb_Albumin +
                    d0_alb_Platelet + d1_alb_Platelet +
                    d0_asl.clin.num_EXINV2 + d1_asl.clin.num_EXINV2 +
                    d0_asl.lab.num_LDH1 + d1_asl.lab.num_LDH1 - 1,
                    data = mydf)

coef(fit)
```

Note:
1. total number of parameters was the same by using R and OTF, means otf fits the same model structure 
2. this also means in the terms results of PfS that we saw before, the intercept was actually fitted, but not shown, while the intercept was actually shown for binary outputs
3. however, parameter values are not the same, although some hints by inverting the signs, means OTF was using different optimizing strategies
4. this is inconvenient, and make it hard for us to debug 
5. key question for the RD team is to interpret the following two parameters: 

-1. Precdictor (ARM = 0)
-2. Precdictor (ARM = 1)

6. especially the second, which could potentially be interpreted in 2 ways:
-1. effects of the predictor when treatment arm equals 1
-2. effects of the predictor when treatment arm equals 1, but conditional on Predictor (ARM = 0)
